### 2021-04-22 input/background.tex

Then, in Figure~\ref{fig:partition_intuitionC}, another \Process~$B$
creates a replica of this content. Similarly to \Process~$R$ and
\Process~$G$, \Process~$B$ can notify other \processes of this new
replica. However, the set of \processes that could actually use this
new replica is actually much smaller than the network size. Uniform
reliable broadcast is not designed for such context and would generate
a lot of unnecessary traffic. Instead, \processes need a communication
primitive that propagates notifications within a scope by evaluating a
predicate, starting at its broadcaster (the \emph{source}).
In other terms, \processes propagate these notifications as long as they
consider them useful based on location information they carry. We call
such primitive \emph{scoped broadcast}, for messages transitively
reach a subset of interconnected \processes (the \emph{scope})
Using this primitive, \processes can lock down the traffic of content
indexing to relevant \processes.

@@

Next Section details the properties of adaptive scoped broadcast in
dynamic networks, and provides an implementation along with its
complexity.

>>>>

Then, in Figure~\ref{fig:partition_intuitionC}, another \Process~$B$
creates a replica of this content. Similarly to \Process~$R$ and
\Process~$G$, \Process~$B$ can notify other \processes of this new
replica. However, the set of \processes that could actually use this
new replica is actually much smaller than the network size. Uniform
reliable broadcast is not designed for such context and would generate
a lot of unnecessary traffic. Instead, \processes need a communication
primitive that propagates notifications within a scope by evaluating a
predicate, starting at its broadcaster (the \emph{source}).  In other
terms, \processes propagate notifications as long as they consider
them useful based on location information they carry. We call such
primitive \emph{scoped broadcast}, for messages transitively reach a
subset of interconnected \processes (the \emph{scope}). Using this
primitive, \processes can lock down the traffic of content indexing to
relevant \processes.

@@

Next Section details the properties of adaptive scoped broadcast in
dynamic networks, and provides an implementation called \NAME along with its
complexity.

### 2021-04-22 input/introduction.tex

In this paper, we first define scoped broadcast as a communication
primitive that allows epidemic dissemination of messages as long as
receiving \processes verify an application-dependant predicate.

\noindent Then, we propose \NAME, a logical partitioning protocol that
uses \underline{s}coped broad\underline{cast} to \underline{a}dapt its
partitions, \ie scopes, to creations and deletions of partitions in
the system. Each source advertises its existence to all possibly
interested \processes, \ie only interconnected \processes of its
partition plus \processes bordering its partition. In turns, \NAME
scales well to the number of sources since the higher the number of
partitions, the earlier the dissemination of notifications stops.
%
%% In this paper, we propose a first implementation of such a protocol.
%% Entitled Adaptive Scoped broadcast (\NAME), 
%% our protocol relies on a primitive that forwards
%% messages until a certain condition is reached, \ie the scope. The
%% scope is defined by a boolean function (predicate) that returns
%% whether a message should be propagated or not around its
%% broadcaster. In the current scenario, the stop condition is related to
%% the distance to the nearest replica : if the message that a node receives
%% indicates a content has a longer distance than that which the node
%% knows then the transfer of the message is useless and so stopped.
%

>>>>

In this paper, we first define scoped broadcast as a communication
primitive that allows epidemic dissemination of messages as long as
receiving \processes verify an application-dependant predicate.  Then,
we propose \emph{\NAME}, a logical partitioning protocol that uses
\underline{s}coped broad\underline{cast} to \underline{a}dapt its
partitions, \ie scopes, to creations and deletions of partitions in
the system. Each source advertises its existence to all possibly
interested \processes, \ie only interconnected \processes of its
partition plus \processes bordering its partition. In turns, \NAME
scales well to the number of sources since the higher the number of
partitions, the earlier the dissemination stops.
%
%% In this paper, we propose a first implementation of such a protocol.
%% Entitled Adaptive Scoped broadcast (\NAME), 
%% our protocol relies on a primitive that forwards
%% messages until a certain condition is reached, \ie the scope. The
%% scope is defined by a boolean function (predicate) that returns
%% whether a message should be propagated or not around its
%% broadcaster. In the current scenario, the stop condition is related to
%% the distance to the nearest replica : if the message that a node receives
%% indicates a content has a longer distance than that which the node
%% knows then the transfer of the message is useless and so stopped.
%

### 2021-04-22 input/proposal.tex

\vfill \TODO{spacing for intro and related work: 2.5 pages}
\newpage

@@

\begin{definition}[Edge infrastructure]
  An edge infrastructure is a connected \underline{g}raph $G(V, E)$ of
  \underline{v}ertices $V$ and bidirectional \underline{e}dges $E
  \subseteq V \times V$.  A \underline{p}ath $\pi_{ij}$ from \Process
  $i$ to \Process $j$ is a sequence of vertices $[i, k_1, k_2, \ldots
    k_n, j]$ with $\forall m: 0\leq m \leq n, \langle \pi_{ij}[m],
  \pi_{ij}[m+1] \rangle \in E$.
\end{definition}

@@

\begin{definition}[\label{def:scoped}\underline{S}coped broad\underline{cast} (\NAMEB)]
  When \Process $x$ scoped \underline{b}roadcasts $b_x(m)$ a
  \underline{m}essage $m$, every correct \process $y$ within a scope
  \underline{r}eceives $r_y(m)$ and \underline{d}elivers it
  $d_y(m)$. The scope depends on the \underline{s}tate $\sigma$ of
  each \process, the \underline{m}etadata $\mu$ piggybacked by each
  message, and a \underline{p}redicate $\phi$ verified from \process to
  \process: {\small{$b_x(m) \wedge r_y(m) \implies \exists \pi_{xy}: \forall z
  \in \pi_{xy}, \phi(\mu_z, \sigma_z)$}}.
\end{definition}

@@

\begin{definition}[\label{def:partitioning}Partitioning]
  Let $S \subseteq V$ be the set of \underline{s}ources, and $P_s$ be
  the \underline{p}artition including at least \Process $s$, each
  \process belongs to at most one partition $\forall p,q \in V, \forall
  s_1,s_2 \in S: p \in P_{s_1} \wedge q \in P_{s_2} \implies p \neq q
  \vee s_1 = s_2$, and there exists at least one path $\pi_{ps}$ of
  \processes that belong to this partition $\forall q \in \pi_{ps}: q
  \in P_s$.
\end{definition}

@@

Unfortunately, \processes do not share a common global knowledge of
the network state. For \processes to reach consistent partitioning
together, a \process becoming a source must notify every \process that
is closer to it than any other source. As stated in
Section~\ref{sec:scoped}, since epidemic dissemination and scoped
broadcast have well-aligned preoccupations, we assume implementations
relying on forwarding of messages from neighbor to neighbor.

\begin{theorem}[\label{theo:bef}\underline{B}est \underline{e}ventual
    \underline{f}orwarding $(BEF) \implies CP$]
%
Assuming reliable communication links where a correct \process $q$
eventually receives the message $m$ \underline{s}ent to it by a
\process $p$ ($s_{pq}(m) \implies r_{q}(m)$); a monotonically growing
system ($G^t \subseteq G^{t+1}$); \processes eventually reach
consistent partitioning if each \process eventually forwards its best
known partition.
\end{theorem}

\begin{proof}
  % \begin{asparadesc}
  % \item [$BEF \implies CP$:]
  When a \process $s_1$ becomes a source, it belongs to its own
  partition, for there exists no better partition than its own:
  $\forall p \in V: |\Pi_{s_1 s_1}| < |\Pi_{s_1 p}|$. It delivers,
  hence forwards such $\alpha$ message to its -- current or future --
  neighbors. Since communication links are reliable, neighboring
  \processes eventually receive such notification $\forall \langle
  s_1, q \rangle \in E, s_1 \in S \iff \eventually
  r_q(\alpha_{s_1}^{w_{s_1 q}})$. Most importantly, whatever the order
  of received messages, every \process $q'$ in this neighborhood --
  such that there exists no better partition $s_2$ than the received
  one -- delivers and forwards it: $\forall s_2 \in S: |\Pi_{q' s_1}|
  < |\Pi_{q' s_2}| \implies \eventually d_{q'}(\alpha_{s_1}^{w_{s_1
      q'}})$. By transitivity, the message originating from $s_1$
  reaches all such \processes through their shortest paths: $\forall
  q'' \in V, s_1, s_2 \in S: |\Pi_{q'' s_1}| < |\Pi_{q'' s_2}|
  \implies \eventually d_{q''}(\alpha_{s_1}^{|\Pi_{s_1 q''}|})$.
  Since there exists only one best sum of weights per \process that
  can never be retracted, \processes eventually reach consistent
  partitioning.
  %% /!\ not equivalence for there exists other implementations.
  %% \item [$CP \implies BEF$:] By contradiction, if a \process $q \in
  %%   \Pi_{sp} = [s, \ldots, q, q', \ldots, p]$ with $s, q, p \in P_s$
  %%   does not forward its received $\alpha_s^{|\Pi_{sq}|}$, then
  %%   following \processes from $q'$ to $p$ may mistake another
  %%   partition for their  because it needs the weight $W_{pq}$.
  %% \end{asparadesc}
\end{proof}

\begin{algorithm}
  \input{input/algoadd.tex}
  \caption{\label{algo:add}Adding a partition by \Process $p$.}
\end{algorithm}

@@

@@

At any time, a source can revoke its self-appointed status of source,
by executing a \texttt{Del} operation, hence deleting its partition
from the system. All \processes that belong to this partition must
eventually choose another partition to belong to. In
Figure~\ref{fig:del}, two partitions exist initially: $P_a$ and $P_d$
that respectively include $\{a\}$, and $\{b, c, d\}$. In
Figure~\ref{fig:delA}, \Process $a$ deletes its partition. It notifies
all neighboring \processes -- here only \Process $b$ -- that may
belong to its partition using \NAMEB. Upon receipt, \Process $b$
discards the message $\delta_a$, for the latter does not target the
former's partition. \Process $b$ sends its own knowledge of best
partition $\alpha_d^3$ that may prove to be the best for \Process
$a$. Eventually, every \processes belong to the same partition
$P_d$. In this scenario, they reach consistent partitioning.

@@

One could guarantee consistent partitioning by always propagating the
$\delta$ messages corresponding to the $\alpha$ messages it propagated
before. In Figure~\ref{fig:problem}, it means that as soon as
\Process~$b$ forwards $\alpha_a^2$, it assumes that its neighbors
\Process~$c$ and \Process~$d$ may need the notification of removal
$\delta_a$ if it exists. However, such solution also implies that
\processes generate traffic not only related to their current
partition, but also related to partitions they belong to in the
past. This would prove overly expensive in dynamic systems where
\processes join or leave the system, create or delete partitions, at
any time.  \TODO{Another paragraph with JJ approach?}  Instead, we
propose to use the delivery order at each \process to detect possible
inconsistencies and solve them. Together, \processes eventually remove
all stale control information (transiting messages and local states)
of the system leaving room for propagation of messages about
up-to-date partitions.

@@

@@

\begin{algorithm}
  \input{input/algoascast.tex}
  \caption{\label{algo:ascast}\NAME at \Process $p$ in static networks.}
\end{algorithm}

@@

\begin{algorithm}
  \input{input/algoedges.tex}
  \caption{\label{algo:edges}\NAME at \Process $p$ in dynamic networks.}
\end{algorithm}

>>>>

% \vfill \TODO{spacing for intro and related work: 2.5 pages}
% \newpage

@@

\begin{definition}[Edge infrastructure]
  An edge infrastructure is a connected \underline{g}raph $G(V, E)$ of
  \underline{v}ertices $V$ and bidirectional \underline{e}dges $E
  \subseteq V \times V$.  A \underline{p}ath $\pi_{ij}$ from
  \Process~$i$ to \Process~$j$ is a sequence of vertices $[i, k_1,
    k_2, \ldots k_n, j]$ with $\forall m: 0\leq m \leq n, \langle
  \pi_{ij}[m], \pi_{ij}[m+1] \rangle \in E$.
\end{definition}

@@

\begin{definition}[\label{def:scoped}\underline{S}coped broad\underline{cast} (\NAMEB)]
  When \Process~$x$ scoped \underline{b}roadcasts $b_x(m)$ a
  \underline{m}essage $m$, every correct \process $y$ within a scope
  \underline{r}eceives $r_y(m)$ and \underline{d}elivers it
  $d_y(m)$. The scope depends on the \underline{s}tate $\sigma$ of
  each \process, the \underline{m}etadata $\mu$ piggybacked by each
  message, and a \underline{p}redicate $\phi$ verified from \process to
  \process: {\small{$b_x(m) \wedge r_y(m) \implies \exists \pi_{xy}: \forall z
  \in \pi_{xy}, \phi(\mu_z, \sigma_z)$}}.
\end{definition}

@@

\begin{definition}[\label{def:partitioning}Partitioning]
  Let $S \subseteq V$ be the set of \underline{s}ources, and $P_s$ be
  the \underline{p}artition including at least \Process~$s$, each
  \process belongs to at most one partition \smash{\small$\forall p,q \in V, \forall
  s_1,s_2 \in S: p \in P_{s_1} \wedge q \in P_{s_2} \implies p \neq q
  \vee s_1 = s_2$}, and there exists at least one path $\pi_{ps}$ of
  \processes that belong to this partition $\forall q \in \pi_{ps}: q
  \in P_s$.
\end{definition}

@@

Unfortunately, \processes do not share a common global knowledge of
the network state. For \processes to reach consistent partitioning
together, a \process becoming a source must notify every \process that
is closer to it than any other source. Since epidemic dissemination
and scoped broadcast have well-aligned preoccupations, we assume
implementations relying on forwarding of messages from neighbor to
neighbor.

\begin{theorem}[\label{theo:bef}\underline{B}est \underline{e}ventual
    \underline{f}orwarding $(BEF) \implies CP$]
%
Assuming reliable communication links where a correct \process $q$
eventually receives the message $m$ \underline{s}ent to it by a
\process $p$ ($s_{pq}(m) \implies r_{q}(m)$), \processes eventually
reach consistent partitioning if each \process eventually forwards its
best known partition.
\end{theorem}

@@

\begin{proof}
  % \begin{asparadesc}
  % \item [$BEF \implies CP$:]
  When a \process $s_1$ becomes a source, it belongs to its own
  partition, for there exists no better partition than its own:
  $\forall p \in V: |\Pi_{s_1 s_1}| < |\Pi_{s_1 p}|$. It delivers,
  hence forwards such $\alpha$ message to its neighbors. Since
  communication links are reliable, neighboring \processes eventually
  receive such notification $\forall \langle s_1, q \rangle \in E, s_1
  \in S \iff \eventually r_q(\alpha_{s_1}^{w_{s_1 q}})$. Most
  importantly, whatever the order of received messages, every \process
  $q'$ in this neighborhood -- such that there exists no better
  partition $s_2$ than the received one -- delivers and forwards it:
  $\forall s_2 \in S: |\Pi_{q' s_1}| < |\Pi_{q' s_2}| \implies
  \eventually d_{q'}(\alpha_{s_1}^{w_{s_1 q'}})$. By transitivity, the
  message originating from $s_1$ reaches all such \processes through
  their shortest paths: $\forall q'' \in V, s_1, s_2 \in S: |\Pi_{q''
    s_1}| < |\Pi_{q'' s_2}| \implies \eventually
  d_{q''}(\alpha_{s_1}^{|\Pi_{s_1 q''}|})$.  Since there exists only
  one best sum of weights per \process that can never be retracted,
  \processes eventually reach consistent partitioning.
  %% /!\ not equivalence for there exists other implementations.
  %% \item [$CP \implies BEF$:] By contradiction, if a \process $q \in
  %%   \Pi_{sp} = [s, \ldots, q, q', \ldots, p]$ with $s, q, p \in P_s$
  %%   does not forward its received $\alpha_s^{|\Pi_{sq}|}$, then
  %%   following \processes from $q'$ to $p$ may mistake another
  %%   partition for their  because it needs the weight $W_{pq}$.
  %% \end{asparadesc}
\end{proof}

\begin{algorithm}
  \input{input/algoadd.tex}
  \caption{\label{algo:add}Adding a partition by \Process~$p$.}
\end{algorithm}

@@

At any time, a source can revoke its self-appointed status of source
by executing a \texttt{Del} operation, hence deleting its partition
from the system. All \processes that belong to this partition must
eventually choose another partition to belong to. In
Figure~\ref{fig:del}, two partitions exist initially: $P_a$ and $P_d$
that respectively include $\{a\}$, and $\{b, c, d\}$. In
Figure~\ref{fig:delA}, \Process~$a$ deletes its partition. It notifies
all neighboring \processes -- here only \Process~$b$ -- that may
belong to its partition using \NAMEB. Upon receipt, \Process~$b$
discards the message $\delta_a$, for the latter does not target the
former's partition. \Process~$b$ sends its own knowledge of best
partition $\alpha_d^3$ that may prove to be the best for
\Process~$a$. Eventually, every \processes belong to the same
partition $P_d$. In this scenario, they reach consistent partitioning.

@@

One could guarantee consistent partitioning by always propagating the
$\delta$ messages corresponding to the $\alpha$ messages it propagated
before. In Figure~\ref{fig:problem}, it means that as soon as
\Process~$b$ forwards $\alpha_a^2$, it assumes that its neighbors
\Process~$c$ and \Process~$d$ may need the notification of removal
$\delta_a$ if it exists. However, such solution also implies that
\processes generate traffic not only related to their current
partition, but also related to partitions they belong to in the
past. This would prove overly expensive in dynamic systems where
\processes join or leave the system, create or delete partitions, at
any time.  Instead, we propose to use the delivery order at each
\process to detect possible inconsistencies and solve them. Together,
\processes eventually remove all stale control information (transiting
messages and local states) of the system leaving room for propagation
of messages about up-to-date partitions.

@@

\vfill

\TODO{spacing.}

\newpage

@@

\begin{algorithm}
  \input{input/algoascast.tex}
  \caption{\label{algo:ascast}\NAME at \Process~$p$ in static networks.}
\end{algorithm}

@@

\begin{algorithm}
  \input{input/algoedges.tex}
  \caption{\label{algo:edges}\NAME at \Process~$p$ in dynamic networks.}
\end{algorithm}
